Nonlinear Observability for Relative Orbit Determination with Angles-Only Measurements \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2018-06-09 Journal/JAS OD EKF/JAS_OD_EKF.tex
journal added

Relative orbit determination (ROD) between spacecraft is desirable for many spacecraft missions. In scenarios where the spacecraft are in close-proximity, this is commonly referred to as relative navigation.
When baselines between the spacecraft are longer, a more appropriate term would be space-based space surveillance or space-based orbit determination. In particular, vision-based ROD has received attention recently, since optical sensors have the desirable properties of being low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles, from a chief to a deputy.

@@

Next, we present a new measure of observability for relative orbit determination with angles-only measurements. The objectives are to understand the degree of observability for the relative orbital dynamics, and to identify the class of relative orbits that have stronger or weaker observability. For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.

@@

In short, the main contributions of this paper are constructing sufficient conditions for the observability of nonlinear relative orbits with angles-only measurements, and proposing a new quantitative measure of observability. The main objective is to provide assurance that relative orbits are observable with angles-only measurements under certain conditions, if the nonlinearity of relative orbital dynamics are properly considered. More comprehensive study for observability of nonlinear relative orbits, such as obtaining necessary conditions for observability or sufficient conditions for unobservability, is deferred for future study.

@@

Observability of nonlinear systems and sufficient conditions for observability are summarized as follows~\cite{HerKreITAC77,Nijvan90}. For a given nonlinear dynamic system described with \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ is called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period. Note that $\x_0$ is indistinguishable from $\x_0$ trivially from the definition. The system is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $V_0$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself. This implies that $\x_0$ is distinguishable from any other points in a certain neighborhood of it.

@@

\paragraph{Observability Matrix}
We apply the above observability criteria for the nonlinear relative orbital dynamics. The first block of rows of the observability matrix is simply given by
\begin{align*}
\deriv{}{\x} L_f^0 h(\x) = \deriv{\y}{\x}.
\end{align*} 
Similarly, from \refeqn{xxdot}, the second row block is
\begin{align*}
\deriv{}{\x} L_f^1 h(\x) = \deriv{}{\x}\parenth{\deriv{\y}{\x}f(\x)}=\deriv{\dot\y}{\x}.
\end{align*} 
Since $\dot\y$ is determined completely from $\x$, we have
\begin{align*}
\deriv{}{\x} L_f^2 h(\x) = \deriv{}{\x}\parenth{\deriv{\dot\y}{\x}f(\x)}=\deriv{\ddot\y}{\x}.
\end{align*}
Using these and the fact that $\x=[\r^T,\dot \r^T]$, the observability matrix of the relative orbital dynamics can be written as
\begin{align}
\mathcal{O} =
\begin{bmatrix}
\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\[0.4em]
\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\[0.4em]
\deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
\end{bmatrix}
\triangleq
\begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix}\in\Re^{9\times 6},\label{eqn:OO}
\end{align}
where the submatrices $\mathcal{O}_{ij}\in\Re^{3\times 3}$ are defined for each block, and we used the fact that the measurement $\y$ is independent of $\dot\r$. Here we consider the observability matrix obtained by up to the second order Lie derivatives of the measurement due to complexity. But, this still provides sufficient conditions for observability under certain geometric conditions described later.

@@

The expression $\r^T(\mathbf{v}_{rel}\times \mathbf{a}_1)$ in the third condition of \refeqn{cond1} corresponds to the (signed) volume of a parallelepiped whose sides are given by the vectors $\r$, $\mathbf{v}_{ref}$, and $\mathbf{a}_1$. Therefore, the third condition is violated if all of the three vectors are coplanar. This may happen when there is no out-of-plane motion of the deputy, i.e., the orbital plane of the deputy is identical to the chief, and $z(t)\equiv 0$ for all $t$. As such, unlike the prior two conditions, the third condition is violated always for any coplanar relative motions. The third condition of \refeqn{cond2} has a similar structure as well.

However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these conditions is not satisfied does not necessarily imply that the considered point is not observable. In such cases, the third or higher order Lie derivatives should be checked to confirm observability. Furthermore, the presented differential geometric approaches to determine observability do not yield necessary conditions for observability. As such, these conditions cannot be used to determine the certain types of relative motions that are not observable with angles-only measurements. This paper is focused on verifying the observability of nonlinear relative orbits via angles-only measurements under certain conditions, and determining observability with higher-order Lie derivatives or necessary conditions for observability is deferred for future investigation.

@@

The observability criteria derived in the previous section guarantee that the nonlinear relative orbital dynamics are indeed observable under certain geometric conditions. However, they are limited in the sense that they only provide binary information about observability. Therefore, they cannot be used to determine whether a certain type of orbit is more easily estimated compared with other orbits based on angles-only measurements. Furthermore, as they are defined in terms of the state vector at a specific time, they do not capture the effects of the measurement trajectory over a certain time period on the observability, which often predicts the performance of estimation.

Motivated by the above issues, we propose a quantitative measure that represents the degree of observability for relative orbits in this section. More explicitly, this measure evaluates the observability in a quantitative manner as it determines how effectively the relative orbit can be estimated by line-of-sight measurements.

The problem of assigning physically meaningful measures of the quality of observability has been introduced in~\cite{BroPNEC66,MonITAC67,MulWebA72} for linear systems, and in particular, in \cite{MulWebA72} the determinant, trace, and maximal eigenvalue of the inverse characteristic observability matrix are maximized with respect to certain parameters of a dynamic system. Several approaches have been proposed for an observability measure of nonlinear systems. 
%The measure of observability has been generalized to nonlinear dynamic systems in order to obtain balanced realizations.
An energy-like observability function is introduced in references~\cite{SchSCL93,Sch94,NewKriPICDC98} to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system. But, this approach is only applicable to asymptotically stable equilibrium states. An empirical observability Gramian is introduced in references~\cite{LalMarPIWC99,LalMarIJRNC02}, which is essentially a covariance matrix of the output computed by a number of sample trajectories.
It has been applied to a balanced realization of chemical processes~\cite{HahEdgJPC03,HahEdgCCE02}. However, this approach is based on the assumption that the output converges to a steady-state value, and it is not clear how sample trajectories are selected. Therefore, these approaches cannot be applied for the relative orbital dynamics that do not have steady-state responses.

@@

In general, the Gramian of a set of time-dependent functions represents the degree of linear independence of those functions over a given period of time. Therefore, the observability Gramian defined in \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent of each other. This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent on $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$. Therefore, the condition number of $\mathcal{W} (\x_0,t_0,t_f)$ can be used as a measure of observability. When applied to linear dynamic systems, we can show that \refeqn{Wo_NL} reduces to the common observability Gramian for linear dynamic systems.

@@

In contrast to the observability criteria presented in Section 3, the proposed observability measure is obtained numerically. As such, it is not useful to identify the particular class of the relative orbits that yield stronger, or weaker observability. However, it can be used to analyze the degree of observability in a complementary manner when the sufficient conditions of the observability criteria are violated, as illustrated by the following numerical examples.

@@

It is assumed that the chief is on a circular orbit with an altitude of $500\,\mathrm{km}$ and zero inclination. Here, we study the effects of the deputy inclination on the observability measure and filtering performance. It is motivated by the fact that the given sufficient conditions for observability are violated if the inclination of the deputy is zero as well, since the deputy is on the same orbital plane with the chief. The eccentricity of the deputy is chosen as $e_{deputy}=0.2$ and the semi-major axis of the deputy is chosen such that its orbital period is identical to the chief, i.e.,
\begin{align*}
a_{deputy} = a_{chief} = \parenth{\frac{T_{chief}\sqrt{\mu}}{2\pi}}^{2/3}.
\end{align*}
Therefore, the relative orbits are periodic. The inclination of the deputy is varied as $i_{deputy}=0,\,10,\,\cdots,\,50^\circ$, where the $i_{deputy}=0$ case corresponds to coplanar orbits such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated.

For the extended Kalman filter, the initial estimate of the state is chosen as two times its true value, i.e., $\hat\x_0=2\x_0$. This implies that there is a large initial error in the magnitude of the state, which is difficult to estimate accurately using angles-only measurements. The initial covariance of the state is chosen as $P_0=\mathrm{diag}[(50\,\mathrm{km})^2I_{3\times 3}, (50n\,\mathrm{km/s})^2I_{3\times 3}]$. The covariance matrices for the process noise and the measurement noise are $Q_k=\mathrm{diag}[10^{-6}I_{3\times 3}, 10^{-8}I_{3\times 3}]$, and \\$R_k=1.306^2 I_{3\times 3}\,\mathrm{deg}^2$. It is assumed that the line-of-sight is measured at every $\Delta t = 0.47$ seconds. It is simulated for 20 orbits of the chief around the Earth. Throughout this paper, the units are chosen as $\mathrm{km}$, $\mathrm{sec}$, and $\mathrm{rad}$, if not specified otherwise.

@@

%\clearpage\newpage%asdf
\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_0.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_0.pdf}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.5\textwidth]{PE_i_0_e.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.5\textwidth]{PE_i_0_xx_norm.pdf}}
}
\caption{Estimation results: $i=0$}\label{fig:i0}
\end{figure}

\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_30.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_30.pdf}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.5\textwidth]{PE_i_30_e.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.5\textwidth]{PE_i_30_xx_norm.pdf}}
}
\caption{Estimation results: $i=30$}\label{fig:i30}
\end{figure}

\begin{figure}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.5\textwidth]{PE_i_30_e_J2.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.5\textwidth]{PE_i_30_xx_norm_J2.pdf}}
}
\caption{Estimation results with $J_2$ perturbations: $i=30$}\label{fig:i30}
\end{figure}

\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_i_50.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.5\textwidth]{OT_traj_PE_i_50.pdf}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.5\textwidth]{PE_i_50_e.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.5\textwidth]{PE_i_50_xx_norm.pdf}}
}
\caption{Estimation results: $i=50$}\label{fig:i50}
\end{figure}
%\begin{figure}
%\centerline{
%	\subfigure[Estimation errors]{
%		\includegraphics[width=0.3\textwidth]{PE_i_50_e_J2.pdf}}
%	\hspace*{0.15\textwidth}
%		\subfigure[Magnitude of state (red:true, blue:estimate)]{
%		\includegraphics[width=0.3\textwidth]{PE_i_50_xx_norm_J2.pdf}}
%}
%\caption{Estimation results with $J_2$ perturbations: $i=50$}\label{fig:i50}
%\end{figure}

%\clearpage\newpage%asdf

@@

Next, we study the effects of varying the size of the relative orbits proportionally on nonlinear observability. As before, the chief is on a circular orbit with the altitude of $500\,\mathrm{km}$ and zero inclination. The initial conditions for the relative position and the velocity are chosen as
\begin{gather*}
\r_0=[r_{x_0},\,0,\,0]^T\,\mathrm{km},\quad
\mathbf{v}_0=[0,\,-2r_{x_0}n,\,0]^T\,\mathrm{km/s},
\end{gather*}
where $r_{x_0}\in\Re$ is the initial relative distance that is varied as $10,\,20,\,50,\,100\,\mathrm{km}$. For the linearized relative orbital dynamics of the two-body problem, these initial conditions yield elliptic orbits around the chief, but for the nonlinear relative orbital dynamics, there is drift proportional to the size of the orbit parameterized by $r_{x_0}$. While the resulting orbits are coplanar such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated, it is still desirable to understand how the orbital size or drift affects nonlinear observability. For the extended Kalman filter, the initial estimate of the state is chosen as $\hat\x_0=1.5\x_0$, and the other parameters are identical to the prior case.

As discussed in Section 3, these relative orbits violate the observability criteria summarized in Proposition 1 since they are coplanar orbits. Therefore, Proposition 1 cannot be applied to investigate the observability of these orbits. Instead, the observability is analyzed by the observability measure proposed in Section 4 as follows.

@@

\afterpage{\clearpage
\begin{figure}[p]
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_20}}
%	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_20}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_20_e.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_20_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=20\,\mathrm{km}$}\label{fig:r0_20}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_50}}
%	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_50}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_50_e.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_50_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=50\,\mathrm{km}$}\label{fig:r0_50}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.5\textwidth]{OT_itraj_PE_r0_JAS_100}}
%	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.5\textwidth]{OT_traj_PE_r0_JAS_100}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_100_e.pdf}}
%	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.5\textwidth]{EKF_PE_r0_100_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=100\,\mathrm{km}$}\label{fig:r0_100}
\end{figure}

@@

It is well known that linearized relative orbital dynamics are not observable with angles-only
measurements. This paper studies the observability for nonlinear relative orbital dynamics, and it derives a set of sufficiency conditions to show that the relative orbit is indeed observable under certain geometric conditions. Also, a quantitative observability measure that represents the degree of observability is proposed. It is illustrated by several numerical examples. Future work objectives include constructing observability criteria based on higher-order Lie derivatives of the measurement, and developing necessary conditions for observability to classify the type of relative orbits that are not observable completely. We also plan to develop nonlinear filtering techniques that can capture the weak coupling between the dynamics of relative motion and the line-of-sight measurements effectively, as well as studying the effects of higher-order gravitational perturbations on the observability.

@@

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics

\bibliography{BibSources}   % name your BibTeX data base

>>>>

Relative orbit determination (ROD) between spacecraft is desirable for many spacecraft missions. In scenarios where the spacecraft are in close-proximity, this is commonly referred to as relative navigation. When baselines between spacecraft are longer, a more appropriate term would be space-based space surveillance or space-based orbit determination. In particular, vision-based ROD has received attention recently, since optical sensors have the desirable properties of being low cost and minimal maintenance, while providing accurate line-of-sight measurements, or angles, from a chief to a deputy.

@@

Next, we present a new measure of observability for relative orbit determination with angles-only measurements. This is motivated to understand the degree of observability for the relative orbital dynamics, and to identify the class of relative orbits that have stronger or weaker observability. For a given initial condition of relative orbits, the proposed observability measure determines how much the corresponding relative orbits are easy or difficult to estimate, thereby providing a quantitative measure for relative orbit determination.

@@

In short, the main contributions of this paper are constructing sufficient conditions for the observability of nonlinear relative orbits with angles-only measurements that have not been studied before, and proposing a new quantitative measure of observability.

@@

Observability of nonlinear systems and sufficient conditions for observability are summarized as follows~\cite{HerKreITAC77,Nijvan90}. For a given nonlinear dynamic system described with \refeqn{xxdot} and \refeqn{y}, a pair of points $\x_0$ and $\x_1$ is called \textit{indistinguishable} if the outputs of the corresponding solutions starting from each of $\x_0$ and $\x_1$ are identical for a certain time period. Note that $\x_0$ is indistinguishable from $\x_0$ trivially from the definition. The system is \textit{locally weakly observable} at $\x_0$, if there exists an open neighborhood $V$ of $\x_0$ such that for every open neighborhood $V_0$ of $\x_0$ contained in $V$, the only indistinguishable point to $\x_0$ is the point $\x_0$ itself. This implies that $\x_0$ is distinguishable from any other points in a neighborhood of it.

@@

\paragraph{Observability Matrix}
We apply the above observability criteria for the nonlinear relative orbital dynamics. The first block of rows of the observability matrix is simply given by
\begin{align*}
\deriv{}{\x} L_f^0 h(\x) = \deriv{\y}{\x}.
\end{align*} 
Similarly, from \refeqn{xxdot}, the second row block is
\begin{align*}
\deriv{}{\x} L_f^1 h(\x) = \deriv{}{\x}\parenth{\deriv{\y}{\x}f(\x)}=\deriv{\dot\y}{\x}.
\end{align*} 
Since $\dot\y$ is determined completely from $\x$, we have
\begin{align*}
\deriv{}{\x} L_f^2 h(\x) = \deriv{}{\x}\parenth{\deriv{\dot\y}{\x}f(\x)}=\deriv{\ddot\y}{\x}.
\end{align*}
Using these and the fact that $\x=[\r^T,\dot \r^T]$, the observability matrix of the relative orbital dynamics can be written as
\begin{align}
\mathcal{O} =
\begin{bmatrix}
\deriv{\y}{\r} & \deriv{\y}{\dot\r}\\
\deriv{\dot\y}{\r} & \deriv{\dot\y}{\dot\r}\\
\deriv{\ddot\y}{\r} & \deriv{\ddot\y}{\dot\r}
\end{bmatrix}
\triangleq
\begin{bmatrix}
\mathcal{O}_{00} & 0_{3\times 3}\\
\mathcal{O}_{10} & \mathcal{O}_{11}\\
\mathcal{O}_{20} & \mathcal{O}_{21}
\end{bmatrix}\in\Re^{9\times 6},\label{eqn:OO}
\end{align}
where the submatrices $\mathcal{O}_{ij}\in\Re^{3\times 3}$ are defined for each block, and we used the fact that the measurement $\y$ is independent of $\dot\r$. Here we consider the observability matrix obtained by up to the second order Lie derivatives of the measurement due to complexity. But, this still provides sufficient conditions for observability under certain geometric conditions described later. %asdf

@@

The expression $\r^T(\mathbf{v}_{rel}\times \mathbf{a}_1)$ in the third condition of \refeqn{cond1} corresponds to the (signed) volume of a parallelepiped whose sides are given by the vectors $\r$, $\mathbf{v}_{ref}$, and $\mathbf{a}_1$. Therefore, the third condition is violated if all of the three vectors are coplanar. This may happen when there is no out-of-plane motion of the deputy, i.e., the orbital plane of the deputy is identical to the chief, and $z(t)\equiv 0$ for all $t$. As such, unlike the prior two conditions, the third condition is violated always for any planar relative motions. The third condition of \refeqn{cond2} has a similar structure as well.

However, \refeqn{cond1} and \refeqn{cond2} are sufficient conditions for observability, and the fact that any of these conditions is not satisfied does not necessarily imply that the considered point is not observable. In such cases, the third or higher order Lie derivatives should be checked to determine observability. This paper is focused on verifying the observability of nonlinear relative orbits via angles-only measurements under certain conditions, and determining observability with higher-order Lie derivatives is referred to future investigation.

@@

The observability criteria derived in the previous section guarantee that the nonlinear relative orbital dynamics are indeed observable under certain geometric conditions. However, it does not provide any information about how strong the observability is.

Several approaches have been proposed for an observability measure of nonlinear systems. 
%The measure of observability has been generalized to nonlinear dynamic systems in order to obtain balanced realizations.
An energy-like observability function is introduced in references~\cite{SchSCL93,Sch94,NewKriPICDC98} to measure the degree of contribution of an initial state to the output, and it is applied to a pendulum system. But, this approach is only applicable to asymptotically stable equilibrium states. An empirical observability Gramian is introduced in references~\cite{LalMarPIWC99,LalMarIJRNC02}, which is essentially a covariance matrix of the output computed by a number of sample trajectories.
It has been applied to a balanced realization of chemical processes~\cite{HahEdgJPC03,HahEdgCCE02}. However, this approach is based on the assumption that the output converges to a steady-state value, and it is not clear how sample trajectories are selected. Therefore, these approaches cannot be applied for the relative orbital dynamics that do not have steady-state responses.

@@

In general, the Gramian of a set of time-dependent functions represents the degree of linear independence of those functions over a given period of time. Therefore, the observability Gramian defined in \refeqn{Wo_NL} measures how much the sensitivities of the output with respect to the initial condition are linearly independent of each other. This represents the degree of observability, since for example, if $\deriv{\y}{\x_{0_i}}$ is linearly dependent to $\deriv{\y}{\x_{0_j}}$, then it is difficult to distinguish the effects on $\delta \x_{0_i}$ to the output $\y$ from the effects of $\delta \x_{0_j}$ on $\y$. Therefore, the condition number of $\mathcal{W} (\x_0,t_0,t_f)$ can be used as a measure of observability. When applied to linear dynamic systems, we can show that \refeqn{Wo_NL} reduces to the common observability Gramian for linear dynamic systems.

@@

@@

It is assumed that the chief is on a circular orbit with an altitude of $500\,\mathrm{km}$ and zero inclination. Here, we study the effects of the deputy inclination on the observability measure and filtering performance. It is motivated by the fact that the given sufficient conditions for observability are violated if the inclination of the deputy is zero as well, since the deputy is on the same orbital plane with the chief. The eccentricity of the deputy is chosen as $e_{deputy}=0.2$ and the semi-major axis of the deputy is chosen such that its orbital period is identical to the chief, i.e.,
\begin{align*}
a_{deputy} = a_{chief} = \parenth{\frac{T_{chief}\sqrt{\mu}}{2\pi}}^{2/3}.
\end{align*}
Therefore, the relative orbits are periodic. The inclination of the deputy is varied as $i_{deputy}=0,\,10,\,\cdots,\,50^\circ$.

For the extended Kalman filter, the initial estimate of the state is chosen as two times its true value, i.e., $\hat\x_0=2\x_0$. This implies that there is a large initial error in the magnitude of the state, which is difficult to estimate accurately using angles-only measurements. The initial covariance of the state is chosen as $P_0=\mathrm{diag}[(50\,\mathrm{km})^2I_{3\times 3}, (50n\,\mathrm{km/s})^2I_{3\times 3}]$. The covariance matrices for the process noise and the measurement noise are $Q_k=\mathrm{diag}[10^{-6}I_{3\times 3}, 10^{-8}I_{3\times 3}]$, and $R_k=1.306^2 I_{3\times 3}\,\mathrm{deg}^2$. It is assumed that the line-of-sight is measured at every $\Delta t = 0.47$ seconds. It is simulated for 20 orbits of the chief around the Earth. Throughout this paper, the units are chosen as $\mathrm{km}$, $\mathrm{sec}$, and $\mathrm{rad}$, if not specified otherwise.

@@

\clearpage\newpage
\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_i_0.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.3\textwidth]{OT_traj_PE_i_0}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{PE_i_0_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{PE_i_0_xx_norm.pdf}}
}
\caption{Estimation results: $i=0$}\label{fig:i0}
\end{figure}

\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_i_30.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.25\textwidth]{OT_traj_PE_i_30.pdf}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{PE_i_30_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{PE_i_30_xx_norm.pdf}}
}
\caption{Estimation results: $i=30$}\label{fig:i30}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{PE_i_30_e_J2.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{PE_i_30_xx_norm_J2.pdf}}
}
\caption{Estimation results with $J_2$ perturbations: $i=30$}\label{fig:i30}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_i_50.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.25\textwidth]{OT_traj_PE_i_50.pdf}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{PE_i_50_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{PE_i_50_xx_norm.pdf}}
}
\caption{Estimation results: $i=50$}\label{fig:i50}
\end{figure}
%\begin{figure}
%\centerline{
%	\subfigure[Estimation errors]{
%		\includegraphics[width=0.3\textwidth]{PE_i_50_e_J2.pdf}}
%	\hspace*{0.15\textwidth}
%		\subfigure[Magnitude of state (red:true, blue:estimate)]{
%		\includegraphics[width=0.3\textwidth]{PE_i_50_xx_norm_J2.pdf}}
%}
%\caption{Estimation results with $J_2$ perturbations: $i=50$}\label{fig:i50}
%\end{figure}

\clearpage\newpage

@@

Next, we study the effects of varying the size of the relative orbits proportionally on nonlinear observability. As before, the chief is on a circular orbit with the altitude of $500\,\mathrm{km}$ and zero inclination. The initial conditions for the relative position and the velocity are chosen as
\begin{gather*}
\r_0=[r_{x_0},\,0,\,0]^T\,\mathrm{km},\quad
\mathbf{v}_0=[0,\,-2r_{x_0}n,\,0]^T\,\mathrm{km/s},
\end{gather*}
where $r_{x_0}\in\Re$ is the initial relative distance that is varied as $10,\,20,\,50,\,100\,\mathrm{km}$. For the linearized relative orbital dynamics of the two-body problem, these initial conditions yield elliptic orbits around the chief, but for the nonlinear relative orbital dynamics, there is drift proportional to the size of the orbit parameterized by $r_{x_0}$. While the resulting orbits are planar such that the third condition of \refeqn{cond1} and \refeqn{cond2} is violated, it is still desirable to understand how the orbital size or drift affects nonlinear observability. For the extended Kalman filter, the initial estimate of the state is chosen as $\hat\x_0=1.5\x_0$, and the other parameters are identical to the prior case.

@@

\afterpage{\clearpage
\begin{figure}[p]
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_r0_JAS_20}}
	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.25\textwidth]{OT_traj_PE_r0_JAS_20}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_20_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_20_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=20\,\mathrm{km}$}\label{fig:r0_20}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_r0_JAS_50}}
	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.25\textwidth]{OT_traj_PE_r0_JAS_50}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_50_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_50_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=50\,\mathrm{km}$}\label{fig:r0_50}
\end{figure}
\begin{figure}
\centerline{
	\subfigure[Orbital trajectory in the inertial frame (red:chief, blue:deputy)]{
		\includegraphics[width=0.3\textwidth]{OT_itraj_PE_r0_JAS_100}}
	\hspace*{0.15\textwidth}
		\subfigure[Orbital trajectory in the LVLH frame]{
		\includegraphics[width=0.25\textwidth]{OT_traj_PE_r0_JAS_100}}
}
\centerline{
	\subfigure[Estimation errors]{
		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_100_e.pdf}}
	\hspace*{0.15\textwidth}
		\subfigure[Magnitude of state (red:true, blue:estimate)]{
		\includegraphics[width=0.3\textwidth]{EKF_PE_r0_100_xx_norm.pdf}}
}
\caption{Estimation results: $r_{x_0}=100\,\mathrm{km}$}\label{fig:r0_100}
\end{figure}

@@

It is well known that linearized relative orbital dynamics are not observable with angles-only
measurements. This paper studies the observability for nonlinear relative orbital dynamics, and it derives a set of sufficiency conditions to show that the relative orbit is indeed observable under certain geometric conditions. Also, a quantitative observability measure that represents the degree of observability is proposed. It is illustrated by several numerical examples. Future works include constructing observability criteria based on higher-order Lie derivatives of the measurement, and developing nonlinear filtering techniques that can capture the weak coupling between the dynamics of relative motion and the line-of-sight measurements effectively, as well as studying the effects of higher-order gravitational perturbations on the observability.

@@

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{BibSources}   % name your BibTeX data base
